---
title: "Gapminder analysis"
author: "Andy Zhang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the first line of my document. This this document I will analyze how life expectancy has changed over time.

# Reading in gapminder data

```{r reading-in-data}
gapminder_df <- read.csv("data/gapminder_data.csv")
```

## A smaller header

### an even smaller header

##### a very small

List:

-   list item 1

-   list item 2

Another type of list:

:   list item 1

:   list item 2

Numbered list:

1.  list item 1

2.  list item 2

    \*Sublist

3.  list item 3

We make text **bold with two asterisks** or *italic with one asterisk*

We can also make insert greek letters $\Delta$ or $\mu$

We can also make [hyperlinks](https://pad.carpentries.org/Spring2024-BIOMI6300)

**This is bold,** this is not

# Calculating summary statistics

```{r calculating-median-pop, echo=FALSE}
med_pop <- median(gapminder_df$pop)

num_obs <- nrow(gapminder_df)
```

Over `r num_obs` observations, the median population was `r round(med_pop/10e6, 3)` million people

# Plotting Life Expectancy Over the Years

```{r line_graph, message = FALSE}
library(ggplot2)
p1 <- ggplot(data = gapminder_df, aes(x= year, y= lifeExp,
                                group = country, color = continent)) +
         labs(x= " Years", 
             y= "Life Expectancy")+
         geom_line(alpha = 0.5)+
         scale_x_log10()+
         scale_y_log10()+
  scale_x_continuous(limits = c(1997, 2007))
p1 + theme_classic()
# Make a line plot
# Log scale y axis
# Bonus: Change the theme
```

# Plotting Population Distribution

```{r histogram, message = FALSE}
ggplot(data = gapminder_df, aes(x = pop)) +
  geom_histogram()
```

# Using visual vs Using Source

| This                | is                         | why                   |
|---------------------|----------------------------|-----------------------|
| Put some stuff here | its a lot easier in visual | Not so easy in source |
|                     |                            |                       |
|                     |                            |                       |

# Adding some new lines because we are working on Git